### 2019-12-31 supplement.Rmd

---
title: "supplmental"
output: html_document
---

## Blood pressure

```{r, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 5 * 1.67, fig.height = 5)

  library(tidyverse)
  source("scripts/functions.R")
  load("scripts/models_july_2019.RData")

  library(sjPlot)

```

Table 3. Mixed-effects model estimates for systolic blood pressure (mmHg)

```{r}

# SBP

  labs <- c("Intercept", "Treatment", "Sex", "Period")

  tab_model(
    me_sbp, me_sbp_p, me_sbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_sbp.html",
    pred.labels = c(labs, "SBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 1. Mixed-effects model estimates for diastolic blood pressure (mmHg)

```{r}

# DBP

  tab_model(
    me_dbp, me_dbp_p, me_dbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_dbp.html",
    pred.labels = c(labs, "DBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )
```

S-Table 2. Mixed-effects model estimates for oscillatory systolic blood pressure (mmHg)
```{r}

# OCS SBP

  tab_model(
    me_osc_sbp, me_osc_sbp_p, me_osc_sbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_osc_sbp.html",
    pred.labels = c(labs, "OSC SBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 3. Mixed-effects model estimates for oscillatory diastolic blood pressure (mmHg)

```{r}

# OCS DBP

  tab_model(
    me_osc_dbp, me_osc_dbp_p, me_osc_dbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_osc_dbp.html",
    pred.labels = c(labs, "OSC DBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 4. Mixed-effects model estimates for central systolic blood pressure (mmHg)

```{r}

# C SBP

  tab_model(
    me_c_sbp, me_c_sbp_p, me_c_sbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_c_sbp.html",
    pred.labels = c(labs, "C SBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 5. Mixed-effects model estimates for central diastolic blood pressure (mmHg)

```{r}

# C DBP

  tab_model(
    me_c_dbp, me_c_dbp_p, me_c_dbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_c_dbp.html",
    pred.labels = c(labs, "C DBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

## Other outcomes

S-Table 6. Mixed-effects model estimates for OXLDL

```{r}
# OXLDL ####

  tab_model(
    me_oxldl, me_oxldl_p, me_oxldl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_oxldl.html",
    pred.labels = c(labs, "OXLDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 7. Mixed-effects model estimates for arterial stiffness

```{r}

# Arterial Stiffness ####

  tab_model(
    me_art_stif, me_art_stif_p, me_art_stif_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_art_stif.html",
    pred.labels = c(labs, "Stiffness Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 8. Mixed-effects model estimates for fasting glucose

```{r}

# Fasting glucose ####

  tab_model(
    me_glu, me_glu_p, me_glu_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_glu.html",
    pred.labels = c(labs, "Glucose Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 9. Mixed-effects model estimates for LDL

```{r}

# LDL ####

  tab_model(
    me_ldl, me_ldl_p, me_ldl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_ldl.html",
    pred.labels = c(labs, "LDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 10. Mixed-effects model estimates for HDL

```{r}

# HDL ####

  tab_model(
    me_hdl, me_hdl_p, me_hdl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_hdl.html",
    pred.labels = c(labs, "HDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 11. Mixed-effects model estimates for total cholesterol

```{r}

# TCHOL ####

  tab_model(
    me_tchol, me_tchol_p, me_tchol_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_tchol.html",
    pred.labels = c(labs, "TCHOL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 12. Mixed-effects model estimates for TAG

```{r}

# TAG ####

  tab_model(
    me_tag, me_tag_p, me_tag_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_tag.html",
    pred.labels = c(labs, "TAG Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

S-Table 13. Mixed-effects model estimates for PWA augment

```{r}

# PWA Augment ####

  tab_model(
    me_pwa_augment, me_pwa_augment_p, me_pwa_augment_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_pwa_augment.html",
    pred.labels = c(labs, "PWA AUGMENT Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```