### 2020-01-01 supplement.Rmd
big update

---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

## Blood pressure

```{r, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 5 * 1.67, fig.height = 5)

# load("scripts/models_dec_2019.RData")
  source("scripts/01_project.R") # Produces the model and data objects

  library(sjPlot)
  library(tidyverse)

```

Mixed-effects model estimates for systolic blood pressure (mmHg)

```{r}

# SBP

  labs <- c("Intercept", "Treatment", "Sex", "Period")

  tab_model(
    me_sbp, me_sbp_p, me_sbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_sbp.html",
    pred.labels = c(labs, "SBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for diastolic blood pressure (mmHg)

```{r}

# DBP

  tab_model(
    me_dbp, me_dbp_p, me_dbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_dbp.html",
    pred.labels = c(labs, "DBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )
```

Mixed-effects model estimates for central systolic blood pressure (mmHg)

```{r}

# C SBP

  tab_model(
    me_c_sbp, me_c_sbp_p, me_c_sbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_c_sbp.html",
    pred.labels = c(labs, "C SBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for central diastolic blood pressure (mmHg)

```{r}

# C DBP

  tab_model(
    me_c_dbp, me_c_dbp_p, me_c_dbp_bl,
    p.val = "kr",
    file = "tables/primary/table_me_c_dbp.html",
    pred.labels = c(labs, "C DBP Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

## Other outcomes

Mixed-effects model estimates for log(OXLDL)

```{r}

# OXLDL ####

  tab_model(
    me_oxldl, me_oxldl_p, me_oxldl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_oxldl.html",
    pred.labels = c(labs, "OXLDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for arterial stiffness

```{r}

# Arterial Stiffness ####

  tab_model(
    me_art_stif, me_art_stif_p, me_art_stif_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_art_stif.html",
    pred.labels = c(labs, "Stiffness Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for fasting glucose

```{r}

# Fasting glucose ####

  tab_model(
    me_glu, me_glu_p, me_glu_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_glu.html",
    pred.labels = c(labs, "Glucose Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for LDL

```{r}

# LDL ####

  tab_model(
    me_ldl, me_ldl_p, me_ldl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_ldl.html",
    pred.labels = c(labs, "LDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for log(HDL)

```{r}

# HDL ####

  tab_model(
    me_hdl, me_hdl_p, me_hdl_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_hdl.html",
    pred.labels = c(labs, "HDL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for log(total cholesterol)

```{r}

# TCHOL ####

  tab_model(
    me_tchol, me_tchol_p, me_tchol_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_tchol.html",
    pred.labels = c(labs, "TCHOL Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for log(TAG)

```{r}

# TAG ####

  tab_model(
    me_tag, me_tag_p, me_tag_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_tag.html",
    pred.labels = c(labs, "TAG Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

Mixed-effects model estimates for PWA augment

```{r}

# PWA Augment ####

  tab_model(
    me_pwa_augment, me_pwa_augment_p, me_pwa_augment_bl,
    p.val = "kr",
    file = "tables/secondary/table_me_pwa_augment.html",
    pred.labels = c(labs, "PWA AUGMENT Baseline"),
    dv.labels = c("Unadjusted", "+ Period effect", "+ Baselines")
    )

```

```{r hdl_variance_plot}

  mutate(hdl, tx = factor(tx, labels = c("Control", "Active"))) %>%
  filter(!is.na(ep)) %>%
  ggplot(aes(x = log(ep), fill = tx)) +
    geom_density(alpha = 0.5, color = "black") +
    theme_minimal() +
    xlab("HDL") +
    ylab("Density") +
    scale_fill_brewer("", palette = "Set1")

```

```{r all_distributions}

  dist(sbp, "SBP (mmHg)", logged = "no")

  dist(dbp, "DBP (mmHg)", logged = "no")

  dist(art_stif, "Arterial Stiffness", logged = "no")

  dist(glu, "Glucose", logged = "no")

  dist(hdl, "HDL", logged = "no")
  dist(hdl, "Log(HDL)", logged = "yes")

  dist(ldl, "LDL", logged = "no")

  dist(oxldl, "OXLDL", logged = "no")
  dist(oxldl, "Log(OXLDL)", logged = "yes")

  dist(pwa_augment, "PWA Augment", logged = "no")

  dist(tag, "TAG", logged = "no")
  dist(tag, "Log(TAG)", logged = "yes")

  dist(tchol, "TCHOL", logged = "no")
  dist(tchol, "Log(TCHOL)", logged = "yes")

```